### 2018-07-30 nips_2018_author_response.tex
re- factored response

\documentclass{article}

\usepackage{nips_2018_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage[rightcaption]{sidecap}
\usepackage{graphicx}

\begin{document}

Something something thank you guys for your rich on point comments that are so
very useful for making science what it is.

\subsubsection*{Choice of original \texttt{DSL}}

As several reviewers pointed out the original \texttt{DSL} influences the efficiency and
accuracy of the algorithm. Reviewer 1 points out ``What happens, for example, if
length is removed [from the original list \texttt{DSL}]?''. To explore that question we
took the original \texttt{Lisp} specification from [citation needed] as well as
tasks from \texttt{Lisp} textbooks to see whether higher level functions would
be discovered. While this example was computationally expensive, it successfully
rediscovered common abstractions from the functional programming world,
including but not limited to \texttt{fold}, \texttt{unfold}, \texttt{map},
\texttt{length}, \texttt{zip}, \texttt{range}, etc.

That being said, the intended goal of the algorithm is not to rediscover
well-known useful abstractions but rather to adapt such abstractions to the
tasks at hands, through more abstractions if need be but not with that as a goal
\textit{per se}.

\begin{SCfigure}[2][h]
  \centering
  \includegraphics[width=7cm]{figures/rebutal/spirals.eps}\label{geomCompiled}
  \caption{Practical example of DSL evolution: originally, the preferred program
  for the leftmost spiral does not use loops while that of the two other does.
  After an iteration the structure of the two rightmost spiral is abstracted.
  For the next iteration the preferred program for the leftmost spiral leverages
  on that and changes in a meaningful way.}
\end{SCfigure}

\subsubsection*{Relationship to Schmid and Kitzelmann}

We thank reviewer three for suggesting additions to the related work. Our
revision will prominently highlight the work of Kitzelmann, Schmid and Katayama,
particularly Kitzelmann and Schmid, whose cognitive inspiration
is closely allied with our motivations.

Kitzelmann and Schmid's \textsc{Igor} systems are sophisticated
analytical inductive programming algorithms that handle certain classes of
problems, much like {Sketch} and {FlashFill} are sophisticated program
synthesizers for certain classes of programs.
We see these lines of work as orthogonal and complementary to that of SCC\@. Our
goal is to take a black-box synthesizer and then learn how to use it to solve a class of
problems. One could replace our enumerative synthesizer
with \textsc{Igor}, Sketch or \textsc{MagicHaskeller}. We chose however to
use a generic and simple enumerative approach because it allowed us to apply SCC
to a diverse spread of problems: functional programming exercises, text editing,
turtle graphics programming, and symbolic regression from images.

This also relates to an insightful
and suggestion for comparison by reviewer 2: FlashFill's DSL is fixed and
manually curated and enhanced for a specific class of problems, over which is
performs well at the loss of agility.

\subsubsection*{More specific questions}

\paragraph*{Reviewer 2: Origin of list task} All the list tasks were manually
created in order to be interesting, challeging, but also human-interpretable for
a prior behaviour experiment. They pre-existed our algorithm.

\paragraph*{Reviewer 2: assesment of the recognition model} The recognition
model's contribution was measured on lesions blabla

\subsubsection*{Reproducibility}
Our source code is publicly available on GitHub,
and a link will be added upon publication (in NIPS or elsewhere),
along with the specific command lines  needed to run
the experiments in the paper.

%% Our project --- including source code, data and automation scripts --- is
%% already available on \url{GitHub} and future revision will include a link to it
%% once the anonymity of the authors can be unveiled. No technical knowledge will
%% be required to reproduced the results.

\end{document}